---
title: "Rolling Window Forward Validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rwFV-protocol}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
    )
```
The function
**stcCropYield::rollingWindowForwardValidation(.)**
implements a variant of the rolling window forward validation protocol
to perform hyperparameter tuning
in order to build a crop yield prediction model.

This vignette gives a brief explanation of the
rolling window forward validation protocol.

### 1. The Rolling Window Forward Validation Protocol

This schematic illustrates a rolling window forward validation scheme
with a **training window** of 5 years, and
a **validation window** of 3 years.
```{r, echo=FALSE, out.width="950px"}
knitr::include_graphics(
    path = file.path("figure-schematic-rwFV.png")
    )
```

Consider the blue box at the bottom, which represents the production cycle 2021.
The five white boxes immediately to its left correspond to the fact that
we are using a training window of 5 years, which means that the training data
for production cycle 2021 will be those from the five years
strictly and immediately prior, namely, 2016 to 2020.

Note that this respects the operational constraint that, for 2021,
the trained prediction model must have been trained on data
from strictly preceding years.
We still need to know which hyperparameter configuration to use
for production cycle 2021, which brings us to these three gray boxes
above the blue box, corresponding to our choice that the validation window
is three years.

The function
**stcCropYield::rollingWindowForwardValidation(.)**
uses simply a grid search for the hyperparameter tuning.
The search grid of hyperparameter configurations is prescribed
by the user at run time.

The rolling window forward validation protocol
is used to choose a particular configuration from
that user-prescribed search grid
to actually use in production for 2021.

Now, fix temporarily an arbitrary candidate hyperparameter configuration from
the user-prescribed search grid.
Using that configuration, we train a model using data
from the following five years: 2015 to 2019.
We then use that resulting trained model to make predictions
for the validation year 2020.
We compute accordingly the parcel-level prediction errors for 2020.

We then aggregate the parcel-level prediction errors down
to a single numerical performance metric.

We repeat for these two other validation years, namely 2018, 2019 .

Averaging the performance metrics across the validation years
-- the grey boxes -- 2018, 2019, 2020,
we arrive finally at a single numerical performance metric / validation error
for the temporarily fixed configuration.

We next repeat for all configurations in our pre-chosen search grid.
And, the optimized configuration to actually be deployed in production
is the one that yields the best aggregated performance metric.

And, that’s it. This is rolling window -- forward validation, or more precisely, our adaptation of it
to the crop yield prediction context.

### 2. Mock Production Error Series (to evaluate the prediction strategy)

One should ask how to tell how well, or how poorly, this proposed strategy
will work in practice.

While we cannot tell what will happen in the future, we can tell EXACTLY
what would have happened in the past, had we deployed this prediction strategy
in the past, which brings us to the orange boxes as shown below.

```{r, echo=FALSE, out.width="950px"}
knitr::include_graphics(
    path = file.path("figure-mock-productions.png")
    )
```

We repeat, for each orange box, what we described above
(namely, executing the rolling window forward validation) for the blue box.
The difference for the orange boxes is the following:
for the blue box, namely the current production year, it is NOT yet possible
to compute the production / prediction errors. But, for the orange boxes, it is.

If we go ahead and do exactly that, that would produce a plot like this:

```{r, echo=FALSE, out.width="950px"}
knitr::include_graphics(
    path = file.path("figure-xgboost_multiphase-mockProdErr-vs-year.png")
    )
```

In the figure above, the orange line illustrates the mock production
error series of the prediction strategy
based on rolling window forward validation.
Each of the gray lines illustrates the validation errors
for a fixed candidate hyperparameter configuration.
